Gammapy: A Python package for gamma-ray astronomy

### 2023-03-21 src/ms-review.tex
further corrections after issue 115
Signed-off-by: Régis Terrier <rterrier@apc.in2p3.fr>

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
%
%
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/irfs.pdf}
	\caption{
		Using \code{gammapy.irf} to read and plot instrument response functions.
		The left panel shows the effective area as a function of energy for
		the \cta, \hess, \magic, \hawc and \fermi instruments. The right panel shows
		the $68\%$ containment radius of the PSF as a function of energy for the \cta, \hess
		and \fermi instruments. The \cta IRFs are from the \textit{prod5} production for the {\it alpha} configuration of the south and north array. The \hess IRFs are from the DL3 DR1,
        using observation ID 033787. The \magic effective area is computed for a
        $20\,{\rm min}$ observation at the Crab Nebula coordinates. The
		\fermi IRFs use \textit{pass8} data and are also taken at the position of the Crab Nebula.
		The \hawc effective area is shown for the event classes $N_{Hit}=5 - 9$ as light gray
		lines along with the sum of all event classes as a black line. The \hawc IRFs are taken from
		the first public release of events data by the \hawc collaboration. All IRFs do not correspond
		to the latest performance of the instruments, but still are representative of the 
		detector type and energy range. We exclusively relied on publicly available
		data provided by the collaborations. The data is also available in the
		\code{gammapy-data} repository.
    }
	\label{fig:irfs}
\end{figure*}
%

The \code{gammapy.irf} sub-package contains all classes and functionalities
to handle IRFs in a variety of formats.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main quantities stored in
the common \gammaray \irfs are the effective area, energy dispersion,
PSF and background rate. The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$), which is defined as the ratio between the
reconstructed energy and the true energy. By definition, the mean of this ratio is
close to unity for a small energy bias and its distribution can 
be typically described by a Gaussian profile. However, more complex
shapes are also common. The migration parameter is given at each offset angle and
reconstructed energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy disperion matrix per sky position. I.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.

>>>>

\subsection{gammapy.irf}
\label{ssec:gammapy-irf}
%
%
\begin{figure*}[t]
	\centering
	\includegraphics[width=1.\textwidth]{figures/irfs.pdf}
	\caption{
		Using \code{gammapy.irf} to read and plot instrument response functions.
		The left panel shows the effective area as a function of energy for
		the \cta, \hess, \magic, \hawc and \fermi instruments. The right panel shows
		the $68\%$ containment radius of the PSF as a function of energy for the \cta, \hess
		and \fermi instruments. The \cta IRFs are from the \textit{prod5} production. The \hess IRFs are from the DL3 DR1,
        using observation ID 033787. The \magic effective area is computed for a
        $20\,{\rm min}$ observation at the Crab Nebula coordinates. The
		\fermi IRFs use \textit{pass8} data and are also taken at the position of the Crab Nebula.
		The \hawc effective area is shown for the event classes $N_{Hit}=5 - 9$ as light gray
		lines along with the sum of all event classes as a black line. The \hawc IRFs are taken from
		the first public release of events data by the \hawc collaboration. All IRFs do not correspond
		to the latest performance of the instruments, but still are representative of the 
		detector type and energy range. We exclusively relied on publicly available
		data provided by the collaborations. The data is also available in the
		\code{gammapy-data} repository.
    }
	\label{fig:irfs}
\end{figure*}
%

The \code{gammapy.irf} sub-package contains all classes and functionalities
to handle IRFs in a variety of functional forms.
Usually, \irfs store instrument properties in the form of multi-dimensional
tables, with quantities expressed in terms of energy (true or reconstructed),
off-axis angles or cartesian detector coordinates. The main quantities stored in
the common \gammaray \irfs are the effective area, energy dispersion,
PSF and background rate. The \code{gammapy.irf}
sub-package can open and access specific \irf extensions,
interpolate and evaluate the quantities of interest on both energy and spatial
axes, convert their format or units, plot or write them into
output files. In the following, we list the main classes of the
sub-package:

@@

\subsubsection{Energy Dispersion}
For \iacts, the energy resolution and bias, sometimes called energy dispersion,
is typically parametrised in terms of the so-called
migration parameter ($\mu$, see section \label{sec:gammaray-data-analysis}).
 However, more complex
shapes are also common. The distribution of the migration parameter is given at each offset angle and
true energy. The main sub-classes are the \code{EnergyDispersion2D} which is
designed to handle the raw instrument description, and the \code{EDispKernelMap},
which contains an energy disperion matrix per sky position, i.e., a 4-dimensional
sky map where each position is associated to an energy dispersion matrix.
The energy dispersion matrix is a representation of the energy resolution
as a function of the true energy only and implemented in \gammapy
by the sub-class \code{EDispKernel}.
